Observation of falling snowflakes orientation at sheltered and unsheltered sites

### 2024-11-29 Review_R1_answers_202412.tex

% LaTeX rebuttal letter example. 
% 
% Copyright 2019 Friedemann Zenke, fzenke.net
%
% Based on examples by Dirk Eddelbuettel, Fran and others from 
% https://tex.stackexchange.com/questions/2317/latex-style-or-macro-for-detailed-response-to-referee-report
% 
% Licensed under cc by-sa 3.0 with attribution required.
% See https://creativecommons.org/licenses/by-sa/3.0/
% and https://stackoverflow.blog/2009/06/25/attribution-required/

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum} % to generate some filler text
\usepackage{fullpage}
\usepackage{xcolor}
\usepackage{float}
\usepackage{soul}
\usepackage{graphicx}
\usepackage{breakcites}
\usepackage{url}

% import Eq and Section references from the main manuscript where needed
% \usepackage{xr}
% \externaldocument{manuscript}

% package needed for optional arguments
\usepackage{xifthen}
\newcommand*{\blue}{\textcolor{blue}}

\begin{document}

\section*{Response to the reviewers of 2024GL113042 }

% General intro text goes here
We thank the reviewers for their critical assessment of our work, their positive feedback and very useful suggestions. 
In the following we address their concerns point by point.\\

\noindent

\begin{table}[H]
\centering
\begin{tabular}{ |c|c| } 
 \hline
 {\itshape italic text  } & {\itshape Comments of the reviewers}  \\ 
 roman text               & Our answers or comments   \\ 
 \blue{blue text}         & \blue{Verbatim quotes from the manuscript}  \\ 
 \hline
\end{tabular}
\end{table}

\noindent
A manuscript is included with highlighted differences and modifications applied during revision is submitted as well.

\subsection*{Editor Comment}
\textit{Thank you for submitting "Observation of falling snowflakes orientation at sheltered and unsheltered sites" [Paper \#2024GL113042] to Geophysical Research Letters. I have received 2 reviews of your manuscript, which are included below and/or attached. As you can see, the reviews indicate that major revisions are needed before we can consider proceeding with your paper. I am therefore returning the paper to you so that you can make the necessary changes.
}

\newpage
\subsection*{Reviewer 1}

\textit{The manuscript is well-written. The authors did a great job addressing the observational issues with the Multi-Angel Snowflake Camera (MASC). I particularly appreciate the reverse-engineering numerical experiments: idealized and realistic objects were generated to evaluate various methods for describing the orientation of those objects. The findings from the numerical experiment, together with the analysis of the in-situ data, will be a valuable addition to the literature and a useful resource for the remote sensing community. I recommend the manuscript be published in GRL, after the following issues are addressed:}

\subsubsection*{Major comments:}
\begin{enumerate}
    \item \textit{The discussion about the impact of the environmental winds on the orientation of the snowflakes is not sound. Wind speed alone may not be sufficient to capture this impact. The distribution of wind components likely plays a more significant role.}
\end{enumerate}

\subsubsection*{Minor comments:}
\begin{enumerate}
    \item \textit{Including results on graupel in a study focused specifically on snowflakes raises questions about the sampling method.}
\end{enumerate}

\clearpage
\subsection*{Reviewer 2}
\textit{The authors present an interesting analysis of the orientation behavior of ice crystals and snowflakes during falling. The analysis is based on a large database of ground-based in-situ observations (Multi-Angle Snowflake Camera). Dependencies of the orientation on observational setups (wind mitigation), environmental parameters and particle properties are discussed. A simulation experiment further reveals important impacts of using different methods to estimate the average orientation angle from multiple cameras.}\\

\textit{Overall, I find the paper very well written and also the non-expert reader can easily follow. I have only a number of minor comments and suggestions. One might of course question whether the topic is of interest to a wider community, as it is a rather specific topic. However, I see the topic itself as highly relevant for active and passive microwave remote sensing but also for estimation of ice particle sedimentation velocity. Therefore, I recommend the manuscript for publication in GRL.}

\subsubsection*{Comments:}

\begin{enumerate}

    \item \textit{There is a very recent laboratory study about ice particle orientation which should be added to reference list and I also highly suggest to discuss their results. They appear to be slightly contradictory to your study:} \cite{Stout_ACP_2024} \\

    We thank the reviewer for this, extremely relevant, recommendation.  This high quality study, with laboratory experiments using snowflake replicas in a liquid solution shows that planar crystals may be falling with regimes that do not have a preferential orientation around 0$^\circ$. We believe that, given some of the assumptions of~\cite{Stout_ACP_2024}, their results are somehow complementary to our work and point out to the need to better understand the role of turbulence in orientation of snow particles. Regarding the cited study, we have the following comments:
    \begin{itemize}
        \item A relevant limitation of this work is that "it  only considers quiescent conditions, as the authors want to know under what conditions particle instability still occurs, even without the addition of turbulence.". This is quite different with respect to our field-based experiments.
        \item The study focuses on planar crystals and it uses 12 sub-types of this  hydrometeors (from plates to dendrites). In our case, the planar crystals are all classified together into an individual class, which is also one of the most difficult to classify with the MASC and to observe (given the 35 microns resolution of the MASC itself). In this sense our study is coarser and cannot go to this level of detail, which could be a reason why we do not observe such behavior. 
        \item Interestingly, another recent study~\cite{Koebschall_EF_2023} using replicas of snow aggregates and a liquid solution, showed instead that the particles tend to orient such that the area projected in the direction of flow is maximized, which is the type of behavior that is consistent with our observations. This is very interesting as two separate indoor studies, with similar methods, seem to indicate that individual crystals of planar type and aggregate snowflakes may be behaving differently regarding preferential orientation.
        \item It is unfortunate that the study reports, as possible backup evidence and as driving motivation, that previous studies using MASC also showed off-zero preferential orientations. This is exactly what we showed, in our paper, to be due to the bias of averaging \textbf{absolute value} of orientation estimates taken from different viewpoints and it was one of the driving motivation of our publication.  
    \end{itemize}

    We now modified the manuscript to give some space to commenting and elaborating on the findings of~\cite{Stout_ACP_2024}. At first, we include this study in Table~1.

    \item \textit{L. 37: Not sure if "standard deviation" is a term that is suited for the plain language summary. Maybe one can formulate simpler: "typical variability of the orientation angle\dots"}\\

    \noindent
    We agree with the reviewer, and we implemented their suggestion. \\

    \item Introduction: Honestly, comparing it with the other sections, I find the introduction a bit weak. Considering the wider audience addressed by GRL, I think the importance of ice particle orientation should be emphasized and explained more. It is certainly true that most radar polarimetric quantities are sensitive to orientation (L. 50-52). But this is by far not all in my opinion. Particle orientation also impacts the scattering signal in passive microwave (MW) observations (both from space and from ground). Many space-borne MW sensors are also polarization-sensitive. Considering the enormous efforts undertaken for example by ECMWF and other services to assimilate those all-sky MW observations, this aspect should be mentioned and cited. Another aspect is that an ice particle falling with a preferred orientation will have a very different sedimentation velocity. So characterizing the particle orientation is quite key to simulate its terminal velocity (which is quite an important parameter in any cloud model). Ice particle orientation is also relevant for determining cloud radiative properties such as optical depth or albedo.

    \item \textit{L. 46: It would be good to add a reference for this statement. I question a bit whether this general statement is actually true. I completely agree for plate-like crystals or large aggregates. But what is for example about bullet rosettes in high cirrus clouds? Or columns/needles? Do we have solid knowledge about their falling orientation? Also the references you provide a few sentences later (\cite{Noel_JAMC_2005,Tinklenberg_JFM_2023}) focus only on planar crystals.}\\

    \item \textit{L. 47: What do you mean by "relatively simple"? Personally, I find all those ice particles terribly complicated.}\\

    In this case our intention was to differentiate between individual crystals (the shape of which can indeed be complicated, but that usually have a clear symmetry) from aggregates and graupel, which are composed by many crystals with additional contribution of riming, making it hard to understand the individual shapes or define a clear geometry. Our phrasing was confusing and this sentence needed to be reformulated in any case (see previous points). It now reads:

     \hl{TO DO}

    \item \textit{Table 1: I like the idea of having a table summarizing previous studies. But are there no studies that looked at columns or needles?}\\

    \cite{Klett_JAS_1995, Reinking_JAMC_1997}

    \item \textit{L. 140: Parenthesis around the reference Leinonen et al., 2021 should be removed.}

    \noindent
    It is now corrected.\\

    \item \textit{L. 157: Why "approximately"?}

    \noindent
    The reviewer is right. The sentence now reads:\\
    \blue{With \textit{Mean\_Abs} the final estimate would be 10$^\circ$, while for \textit{Mean} it would be 3.33$^\circ$}\\

    \item L. 211-215: One general thought regarding the fact that radar studies most often "retrieve" smaller range of orientation angles: If one looks at vertically pointing radar images, one always finds the lowest few hundred meters much more turbulent than the ice and snow cloud aloft. Surface friction is certainly responsible for this effect. So I am wondering if the often observable low turbulence in many ice clouds (for example indicated by small spectrum width) could be the reason why the orientation angle distribution is more narrow? Of course, the discrepancy might also be connected to unrealistic scattering properties often assumed by polarimetric shape retrievals.

    \item Connected to this topic but even more complicated to answer is why some polarimetric radar observations seem to indicate a certain non-zero distribution of the particle orientation: \cite{Melnikov_JAOT_2013}
\end{enumerate}

\bibliographystyle{apalike}
\bibliography{./references.bib}

\end{document}

### 2024-11-29 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

\title{Observation of falling snow orientation in sheltered and unsheltered observation sites}

%\authors{J. Grazioli\affil{1}, M. Condolf\affil{2}, F. Coletti\affil{2}, and A. Berne\affil{1}  }

%\affiliation{1}{Environmental Remote Sensing Laboratory, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland}
%\affiliation{2}{Department of Mechanical and Process Engineering, ETH, Zürich, Switzerland}
% \affiliation{3}{Third Affiliation}

@@

\begin{keypoints}
\item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in non-observational studies. 
\item Sheltered and unsheltered sites show differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
\item Numerical simulations reveal that common estimation choices made in past studies may lead to data processing artifacts.
\end{keypoints}

\begin{abstract}
This study investigates the orientation of falling snowflakes using data from a Multi-Angle Snowflake Camera (MASC). We explore the impact of different observational setups (sheltered vs. unsheltered), wind conditions, hydrometeor type, and axis ratio on the orientation distributions. Numerical simulations are used to select the best orientation estimator and to understand the reason behind different results reported in past literature. We find that previously reported non-zero median orientations are likely artifacts of averaging absolute values of apparent orientations from individual cameras. Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader and flatter distributions observed in unsheltered sites and/or high wind conditions.  Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on snowflake orientation under varying environmental conditions.
\end{abstract}

\section*{Plain Language Summary}
Understanding the orientation of falling snowflakes is important for interpreting weather radar data and better understand snowfall micro-physics. This study uses a 3-view camera to capture detailed images of falling snowflakes and estimate their orientations. Snowflakes generally fall with their largest dimension parallel to the ground (horizontal orientation), with variations in both directions around this setup. We found that observations in sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas. Additionally, our simulations suggest that some previous studies may have incorrectly reported average orientations because of the way data was processed. These findings help refine our knowledge of snowflake behavior and show the need of better understanding the physical mechanisms affecting the distribution of snowflake orientations.

@@

Ice-phase hydrometeors, such as individual crystals and aggregates, fall with a preferred orientation. In stable conditions, they align in a way that maximizes the area projected in the direction of flow \cite{Koebschall_EF_2023}. This fact is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally. The concept of orientation is thus linked to the anisotropic nature of most hydrometeors and loses any meaning for isotropic, spherical particles.

Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values. Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly employ a Gaussian distribution model, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence). Observational studies report as well non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this manuscript, this discrepancy is, in reality, the result of the way multi-view instrumental observations are combined to retrieve an individual estimate of orientation.

The problem of observing hydrometeor orientations while they fall, without generating disturbance induced by the measurement itself, is not trivial. Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}. These relatively recent instruments differ, for example, in the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (on the order of tens of micrometers for MASC and VISSS and hundreds of micrometers for 2DVD and PIP), or instrumental structure (compact for MASC and 2DVD and open-field for PIP and VISSS). These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors. As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.

In this manuscript, we present an analysis of the fall orientation of ice-phase hydrometeors, conducted on a large dataset collected by a MASC. We explore the effect of different siting of the instrument (sheltered or unsheltered), wind conditions, hydrometeor type, and axis ratio, as well as the estimator of orientation used. For the last investigation, we complement the analysis with simulations aimed at uncovering biases that may emerge with some estimation choices.

% ---------------
 \begin{table}
 \caption{Non-exhaustive list of research works that provide assumptions or measurements of the distribution of orientation angles of ice phase particles. $\overline{\theta}$ indicates mean orientation, $\sigma$ is the standard deviation,  $\theta_{M}$ indicates the mode of the distribution, $\theta_{Q50}$ is the median, $MAD(\theta)$ is the median absolute deviation.}
 \label{table:orientations}
 \centering
 \begin{tabular}{l p{80mm}} 
 \hline
  Ref.   & Details \\
 \hline
   \cite{Matrosov_JAM_2001} & $\overline{\theta} = 0^\circ, \sigma = 3^\circ$ \\
   \hline
   \cite{Matrosov_JAS_2005} & $\overline{\theta} = 0^\circ, \sigma = 9^\circ \pm3^\circ$ (dendrites) \\
   \hline
    \cite{Kennedy_JAMC_2011}  &   $\overline{\theta} = 0^\circ, \sigma = 15^\circ$ (dendrites) \newline $\overline{\theta} = 0^\circ, \sigma = 30^\circ$ (aggregates)     \\
   \hline
    \cite{Ryzhkov_JAMC_2011} & $\overline{\theta} = 0^\circ, \sigma = 40^\circ$ (aggregates) \newline 
   $\overline{\theta} = 0^\circ \sigma = 40^\circ$ (graupel) \\
   \hline
   \cite{Hogan_JAMC_2012} & Assumed horizontal  \\
   \hline   
   \cite{Putnam_MWR_2017} & $\overline{\theta} = 0^\circ \sigma = 20^\circ$ (snow) \newline 
   $\overline{\theta} = 0^\circ \sigma \in [0,60]^\circ$ (graupel) \\
   \hline
   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) \newline  $\overline{\theta} = 0^\circ \sigma = 40^\circ$ (aggregates) \\
   \hline
    \cite{Matsui_JGRA_2019}  & $\overline{\theta} = 0^\circ \sigma = 20^\circ$ (aggregates) \newline 
   $\overline{\theta} = 20^\circ \sigma = 42^\circ$ (graupel) \\
   \hline
   \cite{Garrett_GRL_2015}$^*$ & $\theta_{Q50} = 39^\circ, \theta_{M} = 13^\circ$ (aggregates) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 16^\circ$ (rimed particles) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 20^\circ$ (graupel) \newline
   $\theta_{Q50} = 28^\circ, \theta_{M} = 13^\circ$ (low-turbulence, all types) \newline
   $\theta_{Q50}= 40^\circ, \theta_{M} = 23^\circ$ (high-turbulence, all types) \\
   \hline
   \cite{Gergely_JGRA_2016}$^*$ &  $\theta_{Q50} \in [38,47.3]^\circ$, $MAD(\theta) \in [43,51]^\circ$ \\
   \hline
   \cite{Jiang_JAS_2019}$^*$ & $\theta_{M}\approx 10^\circ$ \\
   \hline
   \cite{Fitch_AMT_2021}$^*$ & $\theta_{M} = 57^\circ$ (wind $> 5$ ms$^{-1}$) \newline
   $\theta_{M} = 28^\circ$ (wind $< 1.5$ ms$^{-1}$) \\
   \hline
   \cite{Fitch_JGR_2022}$^*$ & $\theta_{M} = 3^\circ$ 
   \newline
   $\theta_{M} = 13^\circ$ (different estimations)  \\
   \hline
   \cite{Schrom_JAS_2023}$^*$ & Beta distribution (varying parameters) \\
   \hline
    \multicolumn{2}{l}{$^{*}$Papers that present orientation statistics from ground-based measurements.}
 \end{tabular}
 \end{table}

@@

We utilize a large dataset of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}. The MASC instrument~\cite{Garrett_AMT_2012} captures simultaneous images of falling snowflakes with three cameras looking at a common focal point from three different viewpoints (0$^\circ$, +36$^\circ$, -36$^\circ$).

MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet}) and pre-computed geometrical and microphysical descriptors. For the purpose of the present study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, focusing thus on dry snow; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, due to their size being too close to the image resolution (which is about 35$\mu$m); and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}. All filtering steps are conducted on data already integrated into MASCdb, ensuring reproducible results. The resulting filtered dataset comprises 412'150 particles.

@@

In handling three 2D images of an individual 3D object, a primary challenge is to derive a single  estimate of orientation for each snowflake. Snowflakes are commonly modeled as ellipsoids or spheroids in previous MASC studies: the orientation of the major axis of the best-fitting ellipsoid in each 2D image is estimated, and the average of the three views is considered as a proxy for orientation \cite{Garrett_GRL_2015, Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}. The cited studies calculate the mean of the absolute value of the three estimates, an approach referred to here as \textit{Mean\_Abs}. Alternatively, other studies perform a 3D reconstruction of the ellipsoid that best fits all three views and take the orientation of its major axis as the estimate \cite{Jiang_JAS_2019}, a family of approaches named \textit{Ellipsoid\_Fit}. Examining Table~\ref{table:orientations}, the experimental studies using these estimators report the occurrence of non-zero modal orientation values between 10$^\circ$ and 20$^\circ$, challenging the common assumption of a distribution centered at 0$^\circ$. While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach results simply from averaging absolute values. This will offer a reconciliation between the common assumptions made on snowflake orientations and the contradiction of the first studies that used the MASC, a contradiction that was observed also by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their works.

>>>>

\title{Observation of falling snowflakes orientation at sheltered and unsheltered sites}

\authors{J. Grazioli\affil{1}, M. Condolf\affil{2}, Y.-A. Roulet\affil{3} F. Coletti\affil{2}, and A. Berne\affil{1}  }

\affiliation{1}{Environmental Remote Sensing Laboratory, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland}
\affiliation{2}{Department of Mechanical and Process Engineering, ETH, Zürich, Switzerland}
\affiliation{3}{Swiss Federal Office of Meteorology and Climatology (MeteoSwiss)}

@@

\begin{keypoints}
\item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in theoretical studies. 
\item Sheltered and unsheltered sites show systematic differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
\item Numerical simulations reveal that common estimation choices made in past studies may lead to data processing artifacts.
\end{keypoints}

\begin{abstract}
Knowledge about orientation of falling snowflakes is still poorly documented with field measurements despite its importance, for example, in the interpretation of remote sensing data. This study investigates the orientation of  snowflakes using data from a Multi-Angle Snowflake Camera (MASC). We explore the impact of different observational setups (sheltered vs. unsheltered), wind conditions, hydrometeor type, and axis ratio on the orientation distributions. Numerical simulations are used to select the best orientation estimator and to understand the reason behind contrasting results reported in past literature. We find that previously reported non-zero median orientations are likely artifacts due to averaging absolute values of orientations from the three individual cameras. Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader distributions observed at unsheltered sites and/or high wind conditions.  Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on snowflake orientation under varying environmental conditions.
\end{abstract}

\section*{Plain Language Summary}
Understanding the orientation of falling snowflakes is important for interpreting remote sensing data and better understand snowfall micro-physics. This study uses a 3-view camera to capture detailed images of falling snowflakes and estimate their orientations. Snowflakes generally fall with their largest dimension parallel to the ground, with typical variability of the orientation angle of about 30 degrees. We found that observations at sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas. Additionally, our simulations suggest that some previous studies may have incorrectly reported average orientations because of the way data was processed. These findings help refine our understanding of snowflake behavior and show the need to better comprehend the physical mechanisms affecting the distribution of snowflake orientations.

@@

Ice-phase hydrometeors, such as individual crystals and aggregates, are typically assumed to fall with a preferred orientation. It is generally considered that crystals populating frozen clouds aligns preferentially in a way that maximizes the area projected in the vertical direction, as deduced by remote sensing and laboratory experiments. This was considered to be true for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}. 
This assumption is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally. The concept of orientation is thus linked to the anisotropic geometry of most hydrometeors and loses any meaning for isotropic, spherical particles. In the case of snowflakes, they are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.

Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values. Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence). However, they report as well non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this study, this discrepancy is, in reality, the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation.

The problem of observing the orientations of hydrometeors while they fall, without generating disturbance induced by the measurement itself, is not trivial. Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}. These relatively recent instruments differ, for example, in the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (on the order of tens of micrometers for MASC and VISSS and hundreds of micrometers for 2DVD and PIP), or instrumental structure (compact for MASC and 2DVD and open-field for PIP and VISSS). These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors. As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.

In this study, we present an analysis of the fall orientation of ice-phase hydrometeors, conducted on a large dataset collected by a MASC. We explore the effect of sheltered or unsheltered siting of the instrument, wind conditions, hydrometeor type, and axis ratio, as well as the estimator of orientation used. For the last investigation, we complement the analysis with simulations aimed at uncovering biases that may arise with some estimation choices.

% ---------------
 \begin{table}
 \caption{Non-exhaustive list of research works that provide assumptions or measurements of the distribution of orientation angles of ice phase particles. $\overline{\theta}$ indicates mean orientation, $\sigma$ is the standard deviation,  $\theta_{M}$ indicates the mode of the distribution, $\theta_{Q50}$ is the median, $MAD(\theta)$ is the median absolute deviation.}
 \label{table:orientations}
 \centering
 \begin{tabular}{l p{90mm}} 
 \hline
  Ref.   & Details \\
 \hline
   \cite{Matrosov_JAM_2001} & $\overline{\theta} = 0^\circ, \sigma = 3^\circ$ \\
   \hline
   \cite{Matrosov_JAS_2005} & $\overline{\theta} = 0^\circ, \sigma = 9^\circ \pm3^\circ$ (dendrites) \\
   \hline
    \cite{Kennedy_JAMC_2011}  &   $\overline{\theta} = 0^\circ, \sigma = 15^\circ$ (dendrites) \newline $\overline{\theta} = 0^\circ, \sigma = 30^\circ$ (aggregates)     \\
   \hline
    \cite{Ryzhkov_JAMC_2011} & $\overline{\theta} = 0^\circ, \sigma = 40^\circ$ (aggregates or graupel) \\
   \hline
   \cite{Hogan_JAMC_2012} & Assumed horizontal  \\
   \hline   
   \cite{Putnam_MWR_2017} & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (snow) \newline 
   $\overline{\theta} = 0^\circ , \sigma \in [0,60]^\circ$ (graupel) \\
   \hline
   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) \newline  $\overline{\theta} = 0^\circ , \sigma = 40^\circ$ (aggregates) \\
   \hline
    \cite{Matsui_JGRA_2019}  & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (aggregates) \newline 
   $\overline{\theta} = 20^\circ , \sigma = 42^\circ$ (graupel) \\
   \hline
   \cite{Garrett_GRL_2015}$^*$ & $\theta_{Q50} = 39^\circ, \theta_{M} = 13^\circ$ (aggregates) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 16^\circ$ (rimed particles) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 20^\circ$ (graupel) \newline
   $\theta_{Q50} = 28^\circ, \theta_{M} = 13^\circ$ (low-turbulence, all types) \newline
   $\theta_{Q50}= 40^\circ, \theta_{M} = 23^\circ$ (high-turbulence, all types) \\
   \hline
   \cite{Gergely_JGRA_2016}$^*$ &  $\theta_{Q50} \in [38,47.3]^\circ$, $MAD(\theta) \in [43,51]^\circ$ \\
   \hline
   \cite{Stout_ACP_2024}$^*$ &  $\overline{\theta} \in [0,28]^\circ$ (various planar crystals, varying Re number)\\
   \hline
   \cite{Jiang_JAS_2019}$^*$ & $\theta_{M}\approx 10^\circ$ \\
   \hline
   \cite{Fitch_AMT_2021}$^*$ & $\theta_{M} = 57^\circ$ (wind $> 5$ ms$^{-1}$) \newline
   $\theta_{M} = 28^\circ$ (wind $< 1.5$ ms$^{-1}$) \\
   \hline
   \cite{Fitch_JGR_2022}$^*$ & $\theta_{M} = 3^\circ$ 
   \newline
   $\theta_{M} = 13^\circ$ (different estimations)  \\
   \hline
   \cite{Schrom_JAS_2023}$^*$ & Beta distribution (varying parameters) \\
   \hline
    \multicolumn{2}{l}{$^{*}$Papers that present orientation statistics based on field observations.}
 \end{tabular}
 \end{table}

@@

We utilize a large dataset of MASC observations, which includes co-located environmental information such as temperature, humidity, and wind, sourced from the MASCdb database extensively described in \cite{Grazioli_SD_2022}. The database includes observations collected in various sites worldwide, for example in the Swiss Alps, in Korea and in Antarctica between 2014 and 2024. The database includes both observations collected in conditions of natural wind exposure and, for some campaigns, with the instrument installed within a Double-Fence Intercomparison Reference~\cite<DFIR, see for example>{Smith_HESS_2020}, providing enhanced wind sheltering and protection from local turbulence and mean wind .

The MASC instrument captures simultaneous images of falling snowflakes with three coplanar cameras looking at a common focal point from three different viewpoints, with viewing angles of 0$^\circ$, +36$^\circ$, and -36$^\circ$.

MASCdb includes approximately one million image triplets (see Fig.~\ref{fig:triplet} as example) and pre-computed geometrical and microphysical descriptors. For the purpose of the present study, we filter the initial dataset by: (i) excluding precipitation occurring at positive temperatures, as we focus on dry snow and not on melting hydrometeors or rain; (ii) removing particles classified as \textit{small particles} following the criteria in \cite{Praz_AMT_2017}, because of their size being too close to the image resolution (which is about 35$\mu$m) hampering the definition or estimation of their shape and orientation; and (iii) excluding measurements potentially affected or dominated by wind-blown snow according to \cite{Schaer_TC_2020}, as the focus of the study is on precipitation (snowfall) and not horizontal transport. All filtering steps are conducted on data already integrated into MASCdb, ensuring reproducible results. The resulting filtered dataset comprises 412'150 particles.

@@

When handling three 2D images of an individual 3D object, a primary challenge is to derive a single  estimate of orientation for each snowflake. Snowflakes are commonly modeled as ellipsoids or spheroids in previous MASC studies: the orientation of the major axis of the best-fitting ellipsoid in each 2D image is estimated, and the average of the three views is considered as a proxy for orientation \cite{Garrett_GRL_2015, Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}. The cited studies calculate the mean of the absolute value of the three estimates, an approach referred to here as \textit{Mean\_Abs}. Alternatively, other studies perform a 3D reconstruction of the ellipsoid that best fits all three views and take the orientation of its major axis as the estimate \cite{Jiang_JAS_2019}, a family of approaches named \textit{Ellipsoid\_Fit}. Examining Table~\ref{table:orientations}, the experimental studies using these estimators report the occurrence of non-zero modal orientation values between 10$^\circ$ and 20$^\circ$, challenging the common assumption of a distribution centered at 0$^\circ$. While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach simply results from averaging absolute values. This will offer a reconciliation between the common assumptions made on snowflake orientations and the contradiction of the first studies that used the MASC, a contradiction that was also noted by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their work.

To address this, and to select an appropriate orientation estimator, we conducted numerical simulations (results shown in Fig.~\ref{fig:simulations}). These simulations generated ellipsoids with randomly varying geometries, as well as more realistic and geometrically complex aggregates following the same method as \citeA{Leinonen_AMT_2021}. The objects, of known 3D structure are projected into MASC-like 2D images to test various estimation methods. The orientation of the reference snowflake or ellipsoids is known and drawn from a Gaussian distribution with a zero mean and 25$^\circ$ standard deviation. Although this simulation setup could be expanded extensively to investigate the contribution of each ingredient at play, the goal here is only to support the selection of an estimator to use with actual field-collected data. Figure~\ref{fig:simulations} provides an overview of how various methods reproduce the orientation distribution (panels a and c) and the magnitude of errors relative to the absolute orientation value (panels b and d). Tested estimators include: \textit{Min}, representing the minimum (absolute value) orientation among the views with preserved sign; \textit{Mean\_Abs}, as previously defined; \textit{Ellipsoid\_Fit}, as described; \textit{Mean}, the mean of the estimates from the three views; and individual estimates from each view (\textit{Cam0}, \textit{Cam1}, \textit{Cam2}).

\begin{figure}
 \noindent \centering \includegraphics[width=\textwidth]{Fig02.png}
\caption{Evaluation of the retrieval of orientation that could be achieved with a triplet of MASC images, conducted using simulated data (ellipsoids in the left column, simulated snowflakes in the right column). (a) and (c): distribution of orientation value, both reference and reconstructed. (b) and (d): letter-value plots for the  error distribution of the orientation (in absolute value, not sign dependent). The simulations are conducted on 10'000 ellipsoids of various geometry as well as on 10'000 aggregates. }
\label{fig:simulations}
\end{figure}
